Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-05 tests-orszagtang.tex
Updates to sections 11.2.5 through 11.2.10 (minor changes)

\subsubsection{MHD: Orszag-Tang Vortex}
\label{sec.tests.mhd}
Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
classic MHD test problem described by \citep{Orszag79}.  The left panel shows
the result using \enzo's constrained transport MHD method, while the right panel shows the result using
\enzo's implementation of Dedner MHD.
The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
from large scale initial perturbations, and is used to compare the
effective resolution of different MHD schemes.  The test begins with uniform
density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$ (as with
other tests in this section, in the absence of gravity or chemistry we use
arbitrary units).  There is a
single rotational mode in the velocity, and two in the magnetic field:
${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
are accurately represented as compared to, for example,
\citet{Toth00}, and that the resolution of shocks is comparable in
both methods.

>>>>

\subsubsection{MHD: Orszag-Tang Vortex}
\label{sec.tests.mhd}
Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
classic MHD test problem described by \citet{Orszag79}.  The left panel shows
the result using \enzo's constrained transport MHD method, while the right panel shows the result using
\enzo's implementation of Dedner MHD.
The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
from large scale initial perturbations, and is used to compare the
effective resolution of different MHD schemes.  The test begins with uniform
density, $\rho_0=25/(36\pi)$, and pressure, $P_0=5/(12\pi)$ (as with
other tests in this section, in the absence of gravity or chemistry we use
arbitrary units).  There is a
single rotational mode in the velocity, and two in the magnetic field:
${\bf v}_0 = (-\sin(2\pi y) \hat{x},\, \sin(2\pi x) \hat{y})$,
${\bf B}_0 = (-\sin(2\pi y) \hat{x},\, \sin(4\pi x) \hat{y})$.
The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
are accurately represented as compared to, for example,
\citet{Toth00}, and that the resolution of shocks is comparable in
both methods.

### 2013-06-05 tests-pointsourcegravity.tex
Updates to sections 11.2.5 through 11.2.10 (minor changes)

This is a simple test of the ability of the Poisson solver to
correctly reproduce the acceleration around a point source without any
dynamics.  We place a single particle in the center of the unit domain
with root grid dimensions $32^3$ and add a static, nested grid 
refinemed by a factor of two that covers the central $1/8^3$ of the
domain (i.e. so the refined grid is only $8^3$).  We then place 5000
particles throughout the domain distributed uniformly in angle and
in the logarithm of the radius (from the central point).  The
measured accelerations are shown in Figure~\ref{fig.gravitytest}.  As
is typical for Particle-Mesh based methods, the errors decrease at
large distances and peak on roughly the grid scale.  The accelerations
are relatively smooth across the top grid/fine grid boundary with the
most noticeable impact being a drop in the amplitude of tangential
errors on the fine grid.  The force law rolls over at about 1.6 fine
cell widths (equivalent to 0.8 root grid widths in this figure),
consistent with the CIC-deposition and interpolation used.

>>>>

This is a simple test of the ability of the Poisson solver to
correctly reproduce the acceleration around a point source without any
dynamics.  We place a single particle in the center of the unit domain
with root grid dimensions $32^3$ and add a static, nested grid 
refined by a factor of two that covers the central $1/8^3$ of the
domain (i.e. so the refined grid is only $8^3$).  We then place 5000
particles throughout the domain distributed uniformly in angle and
in the logarithm of the radius (from the central point).  The
measured accelerations are shown in Figure~\ref{fig.gravitytest}.  As
is typical for Particle-Mesh based methods, the errors decrease at
large distances and peak on roughly the grid scale.  The accelerations
are relatively smooth across the top grid/fine grid boundary with the
most noticeable impact being a drop in the amplitude of tangential
errors on the fine grid.  The force law rolls over at about 1.6 fine
cell widths (equivalent to 0.8 root grid widths in this figure),
consistent with the CIC-deposition and interpolation used.

### 2013-06-05 tests-zeldovichpancake.tex
Updates to sections 11.2.5 through 11.2.10 (minor changes)

The initial conditions are simple, and we follow the prescription of
\citet{Anninos94}.  Assuming a geometrically flat cosmology, the density
perturbation is given by

\begin{equation}
\rho(x_l) = \rho_0 \big[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \big]
\end{equation}

with the internal energy of the gas set so that the entropy 
is constant throughout.  The velocity perturbation is given by

\begin{equation}
v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l)}{k}
\end{equation}

In the equations above, $\rho_0$ is the background density, z$_c$ is a
free parameter and is the redshift where the sheet forms a caustic
(i.e., where it `pancakes'), z is the redshift of initialization, x$_l$ is the
Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
the perturbation wavelength), and H$_0$ is the value of the Hubble
constant at $z = 0$.  Note that this solution is expressed in terms of 
Lagrangian positions, so one needs to convert this into the Eulerian
coordinates, x$_e$, that are more useful to \enzo:

\begin{equation}
x_e = x_l - \frac{1 + z_c}{1 + z} \frac{\sin(k x_l)}{k}
\end{equation}

We note that the solution described above is exact up to the point of
caustic formation.  In Figure~\ref{fig.pancake}, we show the results
of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
one-dimensional box of length $64$~Mpc/h is initialized at $z = 20$ in
an $\Omega_M = 1$ universe with $h = 0.5$ and a background temperature
of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
grid cells, refining by factors of four using criteria based on cell
mass and the presence of shocks, for a maximum of 2 levels (i.e., an
equivalent maximum resolution of 1024 grid cells).  The simulation is
evolved to z$ = 0$ using the PPM hydro method, with the final output
of the calculation shown in Figure~\ref{fig.pancake}.

Figure~\ref{fig.pancake} shows the key features of this test problem.
The strong shocks and large density gradients are well-resolved, with
density and velocity jumps being well-delineated and at the correct
locations.  The key features of 
this test problem can be resolved with far fewer cells -- simulations
including a mere 8 cells resolve the key features, as shown in
Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
higher resolution here for illustrative purposes.

>>>>

The initial conditions are simple, and we follow the prescription of
\citet{Anninos94}.  Assuming a geometrically flat cosmology, the density
perturbation is given by
\begin{equation}
\rho(x_l) = \rho_0 \left[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \right]
\end{equation}
with the internal energy of the gas set so that the entropy 
is constant throughout.  The velocity perturbation is given by
\begin{equation}
v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l)}{k}
\end{equation}

In the equations above, $\rho_0$ is the background density, z$_c$ is a
free parameter and is the redshift where the sheet forms a caustic
(i.e., where it `pancakes'), z is the redshift of initialization, x$_l$ is the
Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
the perturbation wavelength), and H$_0$ is the value of the Hubble
constant at $z = 0$.  Note that this solution is expressed in terms of 
Lagrangian positions, so one needs to convert this into the Eulerian
coordinates, x$_e$, that are more useful to \enzo:
\begin{equation}
x_e = x_l - \frac{1 + z_c}{1 + z} \frac{\sin(k x_l)}{k}
\end{equation}

We note that the solution described above is exact up to the point of
caustic formation.  In Figure~\ref{fig.pancake}, we show the results
of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
one-dimensional box of length $64$~Mpc/h is initialized at $z = 20$ in
an $\Omega_M = 1$ universe with $h = 0.5$ and a background temperature
of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
grid cells, refining by factors of four using criteria based on cell
mass and the presence of shocks, for a maximum of 2 levels (i.e., an
equivalent maximum resolution of 1024 grid cells).  The simulation is
evolved to z$ = 0$ using the PPM hydro method.

The final output of the calculation, with the key features of this
test problem, is shown in Figure~\ref{fig.pancake}.
The strong shocks and large density gradients are well-resolved, with
density and velocity jumps being well-delineated and at the correct
locations.  The key features of 
this test problem can be resolved with far fewer cells -- simulations
including a mere 8 cells resolve the key features, as shown in
Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
higher resolution here for illustrative purposes.
